---
title: "HW3"
author: "Àngel Casadó (1630811)"
date: "2025-10-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 2

A stock currently trades at \$60. A Call with strike price \$58 and expiry 12 months trades at \$3; a Put with the same strike and expiry trades at \$2. The interest rate at 12 months is 10\%.

-Do any arbitrage opportunities exist?

```{r}
C <- 3
P <- 2
K <- 58
S0 <- 60
t <- 0
r <- 0.10
T <- 1

K_disc <- K * exp(-r * (T-t))
left_side <- C + K_disc
right_side <- P + S0
diff<-right_side-left_side
diff<-round(diff,2)
list(left_side = left_side, right_side = right_side, diff=diff)
```

We can observe how this equality $P+S_0 \ \ \neq \ \ C+Ke^{-r*(T-t)}$, does not hold, and therefore the put-call parity indicates that there is indeed an arbitrage opportunity.

-If there is a possible arbitrage, then explain the strategy to get profit from it.

The difference between both sides of the put–call parity equation represents the the arbitrage profit, which is $`r diff`.

The strategy said that if $S_T>K$ we use CALL to buy without additional cost, but if $S_T<K$ PUT paid the K, closing the operation.

In both cases, there are no further cash flows at maturity, so the profit of $`r diff` is risk-free and requires no forward contract.

